Software Citation Principles

### 2016-03-08 sc-principles/software-citation-principles.tex
broke 5 into 5 (discussion) and 6 (implementation examples) started writing in 5, on derived software

\begin{enumerate}
\item \textbf{Importance}:
Software should be considered a legitimate, citable product of research. Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data. Software should be cited whenever and wherever a research product (such as a paper or derived software) relies upon it, specifically, as part of the standard reference list for that research product. 
\item \textbf{Credit and Attribution}:
Software citations should facilitate giving scholarly credit and normative
and legal attribution to all contributors to the software, recognizing
that a single style or mechanism of attribution may not be applicable to
all software. 
\item \textbf{Unique Identification}:
A software citation should include a method for identification that is
machine actionable, globally unique, interoperable, and recognized by
a community.
\item \textbf{Persistence}:
Unique identifiers and metadata describing the software and its disposition should persist~-- even beyond the lifespan of the software they describe.
\item \textbf{Accessibility}:
Software citations should permit and facilitate access to the software itself and to such associated metadata, documentation, data, and other materials as are necessary for both humans and machines to make informed use of the referenced software. 
\item \textbf{Specificity}:
Software citations should facilitate identification of, and access to, the specific version of software that was used.  Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
%\item \textbf{Interoperability and Flexibility}:
%Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
\end{enumerate}

@@

In the next section (\S\ref{sec:use_cases}), we provide some detailed context in which
software citation is important, by means of use cases.  In \S\ref{sec:related_work}, we
summarize and analyze a large amount of previous work and thinking in this area.  In
\S\ref{sec:principles}, we provide a set of guiding principles for citation of software within
scholarly literature, other software, datasets, or any other research object.  Finally,
in \S\ref{sec:examples} we give examples of how these software citation principles
could be applied, related back to the use cases in \S\ref{sec:use_cases}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion and implementation examples}
\label{sec:examples}

@@

\paragraph{importance}

\katznote{what about derived software?  What metadata should be preserved from the original? What is done for data today?  should this be in the principles?  tried to add to principle 1 and can discuss in next section. maybe also discuss somewhere else, such as in the list of stuff we recognize but don't address (currently in a GitHub issue)}
\LJHnote{this is analogous to built upon research}

@@

\subsection{Implementation examples}

>>>>

\begin{enumerate}
\item \textbf{Importance}: \label{principle:importance}
Software should be considered a legitimate, citable product of research. Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data. Software should be cited whenever and wherever a research product (such as a paper or derived software) relies upon it, specifically, as part of the standard reference list for that research product. 
\item \textbf{Credit and Attribution}:
Software citations should facilitate giving scholarly credit and normative
and legal attribution to all contributors to the software, recognizing
that a single style or mechanism of attribution may not be applicable to
all software. 
\item \textbf{Unique Identification}:
A software citation should include a method for identification that is
machine actionable, globally unique, interoperable, and recognized by
a community.
\item \textbf{Persistence}:
Unique identifiers and metadata describing the software and its disposition should persist~-- even beyond the lifespan of the software they describe.
\item \textbf{Accessibility}:
Software citations should permit and facilitate access to the software itself and to such associated metadata, documentation, data, and other materials as are necessary for both humans and machines to make informed use of the referenced software. 
\item \textbf{Specificity}:
Software citations should facilitate identification of, and access to, the specific version of software that was used.  Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
%\item \textbf{Interoperability and Flexibility}:
%Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
\end{enumerate}

@@

In the next section (\S\ref{sec:use_cases}), we provide some detailed context in which
software citation is important, by means of use cases.  In \S\ref{sec:related_work}, we
summarize and analyze a large amount of previous work and thinking in this area.  In \S\ref{sec:discussion}, we discuss issues related to the principles stated in \S\ref{sec:principles}, and finally,
in \S\ref{sec:examples} we give examples of how these software citation principles
could be applied, related back to the use cases in \S\ref{sec:use_cases}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
\label{sec:discussion}

In this section, we discuss some the principles stated in Section~\ref{sec:principles} and some related concerns.

@@

\paragraph{Derived Software}

The goals of software include the linked ideas of crediting those responsible for software and understanding the dependencies of scientific products on specific software.
In the Importance principle (\ref{principle:importance}), we state that
``software should be cited whenever and wherever a research product (such as a paper or derived software) relies upon it, specifically, as part of the standard reference list for that research product.''
In the case of one code that is derived from another code, citing the derived software may appear to not credit those responsible for the original software, nor recognize its role in the work that used the derived software.
However, this is really analogous to how any research builds on other research, where each research product just cites those products that it directly builds on, not those that it indirectly builds on.
Understanding these chains of knowledge and credit have been part of history of science field for some time, though more recent work is suggesting more nuanced evaluation of the credit chains~\cite{casrai-credit, transitive_credit_json-ld}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation examples}
\label{sec:examples}
